---
title: "Introduction to riskscorer"
author: "Alexander Meyer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to riskscorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, results='hide', echo=FALSE}
library(pander)
library(printr)
library(wakefield)
library(dplyr)

knitr::opts_chunk$set(dev = "CairoPNG")

```


The package *riskscorer* is an attempt to provide an easy programatical interface to clinical risk models. It supports fuzzy argument matching and is webservice ready via [plumber](http://plumber.trestletech.com/) decoration.
A the moment only operative risk scores in cardiac surgery are implemented:

* [STS Score](http://riskcalc.sts.org/)
* [EuroScore I](http://www.euroscore.org)
* [EuroScore II](http://www.euroscore.org)

The programatical interface can be used in various scenarios:
* Simulation research based on a risk score model (see Figure 1)
* Data pre-processing before analysis.
* Usage of the webservice for rapid batch processing of a data pool

Figure 1 shows a simulated data set. Except age and sex, all covariates are fixed.
  > **Crea: 0.8 - 3.0 mg/dL**
  > **Sex: Male or Female**
  > Procedure: CABG+AVR
  > Weight: 75 kg
  > Height: 170 cm
  > Symptoms of heart failure
  > Age: 75
  > LVEF: 60 %
  
```{r data-sim, cache=TRUE}
# n = 1
# df_sim <- r_data_frame(
#     n = n,
#     id,
#     race(x = c("White", "Hispanic", "Black", "Asian"), prob = c(0.8, 0.1, 0.05, 0.05), name = "Race"),
#     age(x = 50:90),
#     nyha = level,
#     sex,
#     lvef = round(rnorm(n = n, mean = 50, sd = 10), digits = 0),
#     crea = round(rnorm(n = n, mean = 1.1, sd = 0.2), digits = 3),
#     height_cm,
#     group(x = c("Yes", "No"), prob = c(1, 0), name = "CABG"),
#     group(x = c("AVR", "MVReplacement"), prob = c(0.7, 0.3), name = "VALVE")
# )
# df_sim

df_sim <- dplyr::data_frame(
  crea = rep(seq(0.8, 3, length.out = 45), times = 2),
  gender = rep(c("Male", "Female"), each = 45)
)
df_sim$lvef = 60
df_sim$age = 75
df_sim$proc_cabg = TRUE
df_sim$proc_valve = "AVR"
df_sim$chf_2w = TRUE
df_sim$height_cm = 170
df_sim$weight_kg = 75

sims <- df_sim %>% group_by(crea, gender) %>% do(
  calc_sts(proc_valve = .$proc_valve, 
           proc_cabg = .$proc_cabg, 
           age = .$age, 
           gender = .$gender,
           chf_2w = .$chf_2w,
           crea = .$crea,
           height_cm = .$height_cm,
           weight_kg = .$weight_kg,
           lvef = .$lvef)
  #data_frame(proc_valve = .$VALVE, proc_cabg = .$CABG, age = .$Age, gender = .$Sex, crea = .$crea, lvef = .$lvef)
)

sims_long <- tidyr::gather(data = sims, key = Risk, value = probability, Mortality:Short_LOS)

ggplot2::qplot(x = crea, y = probability, facets = ~gender, color = Risk, data = sims_long, geom = c("line")) + ggplot2::theme_bw()

ggplot2::qplot(x = crea, y = probability, facets = ~Risk, color = gender, data = sims_long, geom = c("line")) + ggthemes::theme_stata()
```


## STS Score Implemtation

Describe approach of accessing http://riskcalc.sts.org/stswebriskcalc/v1/calculate/stsall
via an json interface and fuzzy argument matching and translation for easy usage.

Works well witg pipes.

Show a generated figure

Talk about Terms of usage

Present web service posibbility with plumber --> potential for automatation of risk score calculation for databases/PDMS systems

## ES I and ES II implementation

Terms and conditions disalllow usage or reverse engenierring.

Risk model is simpler and coefficients are published in [XYZ et al., 20XX](http://citation)
- ES I
- ES II 

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.



## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
