---
title: "Introduction to riskscorer"
author: "Alexander Meyer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to riskscorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, results='hide', echo=FALSE}
library(pander)
library(printr)
library(wakefield)

knitr::opts_chunk$set(dev = "CairoPNG")

```


The package riskscorer is an attempt to provide an easy interface to clinical risk models. A the moment only operative risk scores in cardiac surgery are implemented:

* [STS Score](http://riskcalc.sts.org/)
* [EuroScore I](http://www.euroscore.org)
* [EuroScore II](http://www.euroscore.org)

## Generating a patient data set

```{r data-sim, cache=TRUE}
library(dplyr)

n = 1
df_sim <- r_data_frame(
    n = n,
    id,
    race(x = c("White", "Hispanic", "Black", "Asian"), prob = c(0.8, 0.1, 0.05, 0.05), name = "Race"),
    age(x = 30:90),
    nyha = level,
    sex,
    lvef = round(rnorm(n = n, mean = 50, sd = 10), digits = 0),
    crea = round(rnorm(n = n, mean = 1.1, sd = 0.2), digits = 3),
    height_cm,
    group(x = c("Yes", "No"), prob = c(1, 0), name = "CABG"),
    group(x = c("AVR", "MVReplacement"), prob = c(0.7, 0.3), name = "VALVE")
)
df_sim

df_sim[1:50,] = df_sim[1, ]

df_sim$Age = 31:80

sims <- df_sim %>% group_by(Age) %>% do(
  calc_sts(proc_valve = .$VALVE, proc_cabg = .$CABG, age = .$Age, gender = .$Sex, crea = .$crea, lvef = .$lvef)
  #data_frame(proc_valve = .$VALVE, proc_cabg = .$CABG, age = .$Age, gender = .$Sex, crea = .$crea, lvef = .$lvef)
)

sims_long <- tidyr::gather(data = sims, key = category, value = probability, mort:short_los)

ggplot2::qplot(x = Age, y = probability, color = category, data = sims_long, geom = c("line")) + ggplot2::theme_bw()
```


## STS Score Implemtation

Describe approach of accessing http://riskcalc.sts.org/stswebriskcalc/v1/calculate/stsall
via an json interface and fuzzy argument matching and translation for easy usage.

Works well witg pipes.

Show a generated figure

Talk about Terms of usage

Present web service posibbility with plumber --> potential for automatation of risk score calculation for databases/PDMS systems

## ES I and ES II implementation

Terms and conditions disalllow usage or reverse engenierring.

Risk model is simpler and coefficients are published in [XYZ et al., 20XX](http://citation)
- ES I
- ES II 

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Pandoc tester
```{r,warning=FALSE,message=FALSE}
pander::Pandoc.brew(text = "This is a test: <%= runif(10)%>")
pander::Pandoc.brew(text = "This is a test: <%= as.list(runif(10))%>")
pander::Pandoc.brew(text = "This is a test: \n\n<%= mtcars[1:3, ]%>")
#pander::Pandoc.brew(text = "This is a test: \n\n<%= plot(1:10)%>")
pander::Pandoc.brew(text='<%for (i in 1:5) {%>\n Pi has a lot (<%=i%>) of power:<%=pi^i%><%}%>')

```


## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
