---
title: "Introduction to riskscorer"
author: "Alexander Meyer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The package riskscorer is an attempt to provide an easy interface to clinical risk models. A the moment only operative risk scores in cardiac surgery are implemented:
- [STS Score](http://riskcalc.sts.org/)
- [EuroScore I](http://www.euroscore.org)
- [EuroScore II](http://www.euroscore.org)

## Generating a patient data set

> use the package wakefield here

## STS Score Implemtation

Describe approach of accessing http://riskcalc.sts.org/stswebriskcalc/v1/calculate/stsall
via an json interface and fuzzy argument matching and translation for easy usage.

Works well witg pipes.

Show a generated figure

Talk about Terms of usage

Present web service posibbility with plumber --> potential for automatation of risk score calculation for databases/PDMS systems

## ES I and ES II implementation

Terms and conditions disalllow usage or reverse engenierring

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
