---
title: "Introduction to _riskscorer_"
author: "Alexander Meyer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to riskscorer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, results='hide', echo=FALSE, cache=FALSE, cache.vars=TRUE}
#library(pander)
#library(printr)
#library(wakefield)
suppressPackageStartupMessages(suppressMessages(suppressWarnings({
  library(dplyr)
  library(ggplot2)
  library(riskscorer)
  library(wakefield)
  library(threejs)
})))

knitr::opts_chunk$set(dev = "CairoPNG",
                      fig.width = 6,
                      fig.height = 5,
                      cache=FALSE,
                      echo = FALSE)



```


The package *riskscorer* is a set of functions to provide an easy programmatical interface to clinical risk models. It supports fuzzy argument matching and is webservice ready via [plumber](http://plumber.trestletech.com/) decoration.
A the moment two operative risk scores in cardiac surgery are implemented:

* [STS Score](http://riskcalc.sts.org/)
* [EuroScore II](http://www.euroscore.org)

The following less used scores are under development:

* [EuroScore I](http://www.euroscore.org)
* [German AV Score](http://doi.org/10.1093/ejcts/ezt114)

The programmatical interface can be used in various scenarios:

1. Simulation research based on a risk score model (see Figure 1)
2. Comparison of risk models
3. Data pre-processing and acurate risk score gathering before analysis. (saves work at data entry stage, ensures reproduceibility and thus avoids errors)
4. Usage of the webservice deployment
    + for rapid batch processing of a data pool
    + for clinical information systems to request on demand score calculation

### Example 1: Simulated patient data. Except creatinine, LV function and gender, all covariates are fixed.

The simulated patients have the following characteristics:

* __Creatinine: 0.8 - 3.0 mg/dL__
* __Left ventricular function: 10 - 60 %__
* __Sex: Male or Female__
* Procedure: CABG(Bypass surgery) + AVR(aortic valve replacement)
* Weight: 75 kg
* Height: 170 cm
* Symptoms of heart failure
* Age: 75

The graphs plot the probability of 30d mortality against the left ventricular function and  creatinine value.
  
```{r data-sim-3d}
n = 5#23

grid_df <- expand.grid(crea = rep(seq(0.8, 3, length.out = n)),
            lvef = rep(seq(10, 60, length.out = n)),
            gender = c("Male", "Female"))

df_sim <- grid_df %>% as.data.frame %>% as.tbl

df_sim$age = 75
df_sim$proc_cabg = TRUE
df_sim$proc_valve = "AVR"
df_sim$chf_2w = TRUE
df_sim$height_cm = 170
df_sim$weight_kg = 75

sims <- df_sim %>% group_by(lvef, crea, gender) %>% do(
  calc_sts(proc_valve = .$proc_valve, 
           proc_cabg = .$proc_cabg, 
           age = .$age, 
           gender = .$gender,
           chf_2w = .$chf_2w,
           crea = .$crea,
           height_cm = .$height_cm,
           weight_kg = .$weight_kg,
           lvef = .$lvef)
  #data_frame(proc_valve = .$VALVE, proc_cabg = .$CABG, age = .$Age, gender = .$Sex, crea = .$crea, lvef = .$lvef)
)

sims_long <- tidyr::gather(data = sims, key = Risk, value = probability, Mortality:Short_LOS)

prom <- sims_long %>% dplyr::filter(Risk == "Mortality")
threejs::scatterplot3js(y = prom$crea, x= prom$lvef, z = prom$probability,
                        color = ifelse(prom$gender == "Male", "lightblue", "red"),
                        grid ="TRUE",
                        labels=sprintf("Sex=%s, Crea=%.2f mg/dL, LVEF=%.2f %%, PROM=%.2f %%", 
                                       prom$gender,
                                       prom$crea, 
                                       prom$lvef, 
                                       prom$probability*100),
                        renderer = "canvas")


#plotly::3d

# plot3D::scatter3D(y = prom$crea, x= prom$lvef, z = prom$probability)
# plot3D::scatter3D(y = prom$crea, x= prom$lvef, z = prom$probability,
#                   phi = 40, theta = 70)
# plot3D::scatter3D(y = prom$crea, x= prom$lvef, z = prom$probability,
#                   phi = 0, theta = 70)
# plot3D::scatter3D(y = prom$crea, x= prom$lvef, z = prom$probability,
#                   phi = -20, theta = 70)
# plot3D::scatter3D(y = prom$crea, x= prom$lvef, z = prom$probability,
#                   phi = -30, theta = 70)


# prom_gen <- prom %>% dplyr::filter(gender == "Female")
# plot3D::scatter3D(y = prom_gen$crea, x= prom_gen$lvef, z = prom_gen$probability,
#                   theta = 50, phi = 10, ticktype = "detailed", bty = "f",
#                   zlim = c(0, 0.1), d=2, pch=1,
#                   main ="STS Scores",
#                   xlab = "LVEF [%]",
#                   ylab = "Crea [mg/dL]",
#                   zlab = "STS PROM %",
#                   clab = "STS PROM %")
# prom_gen <- prom %>% dplyr::filter(gender == "Male")
# plot3D::scatter3D(y = prom_gen$crea, x= prom_gen$lvef, z = prom_gen$probability,
#                   theta = 50, phi = 10, pch = 16, add = TRUE, colkey = FALSE)

# library(ggplot2)
# ggplot(prom, aes(x = lvef, y = crea, z = probability, fill = probability)) +
#   facet_wrap(~ gender) + 
#   geom_raster() +
#   scale_fill_gradient("STS PROM %", low = "red", high = "yellow") + 
#   geom_contour(color = "black", bins = 20) + 
#   labs(x = "LVEF [%]", y = "Crea [mg/dL]") +
#   theme_minimal()
# 
# 
# library(lattice)
# lattice::wireframe(probability ~ lvef + crea, data = prom,
#                    groups = gender,
#                    drape = TRUE,
#                    colorkey = TRUE,
#                    shade = FALSE,
#                    xlab = "LVEF [%]",
#                    ylab = "Crea [mg/dL]",
#                    zlab = "STS PROM %")
# 
# scatterplot3d::scatterplot3d(y = prom$crea,
#                              x= prom$lvef, 
#                              z = prom$probability, 
#                              #highlight.3d=TRUE, 
#                              grid = TRUE, pch=16,
#                              box = FALSE,#type = "l",
#                              color = ifelse(prom$gender == "Male", "blue", "red"))
```


### Example 2: Simulated patient data. Except creatinine and gender, all covariates are fixed.

The simulated patients have the following characteristics:

* __Creatinine: 0.8 - 3.0 mg/dL__
* __Sex: Male or Female__
* Left ventricular function: 45 %
* Procedure: AVR(aortic valve replacement)
* Weight: 75 kg
* Height: 170 cm
* Symptoms of heart failure
* Age: 75

The graphs plot the probability of each risk category against the gender and creatinine value.

```{r data-sim-2d}
n = 3#23

grid_df <- expand.grid(crea = rep(seq(0.8, 3, length.out = n)),
            gender = c("Male", "Female"))

df_sim <- grid_df %>% as.data.frame %>% as.tbl

df_sim$age = 75
df_sim$lvef = 45
df_sim$proc_cabg = TRUE
df_sim$proc_valve = "AVR"
df_sim$chf_2w = TRUE
df_sim$height_cm = 170
df_sim$weight_kg = 75

sims <- df_sim %>% group_by(crea, gender) %>% do(
  calc_sts(proc_valve = .$proc_valve, 
           proc_cabg = .$proc_cabg, 
           age = .$age, 
           gender = .$gender,
           chf_2w = .$chf_2w,
           crea = .$crea,
           height_cm = .$height_cm,
           weight_kg = .$weight_kg,
           lvef = .$lvef)
)

sims_long <- tidyr::gather(data = sims, key = Risk, value = probability, Mortality:Short_LOS)

qplot(x = crea, y = probability, facets = ~gender, color = Risk, data = sims_long, geom = c("line")) + theme_bw()

qplot(x = crea, y = probability, facets = ~Risk, color = gender, data = sims_long, geom = c("line")) + theme_bw()

```

### Example 3: clusters of a two cohorts of typical multimorbid TAVR (transcatheter heart valve replacement) and typical younger CABG only patients

The simulated patients have the following characteristics:

* TAVI cohort
    + Procedure: AVR(aortic valve replacement)
    + Gender: : random distribution (probability: M = 0.5122, no = 0.4878)
    + Age: random age in range __78 - 100__ years
    + Left ventricular function: random lvef (mean=50, sd=10)
    + Creatinine: random creatinine value (__mean=1.2, sd=0.2__)
    + Symptoms of heart failure: random distribution (probability: __yes = 0.9, no = 0.1__)
* CABG cohort
    + Procedure: CABG
    + Gender: : random distribution (probability: M = 0.5122, no = 0.4878)
    + Age: random age in range __40 - 70__ years
    + Left ventricular function: random lvef (mean=50, sd=10)
    + Creatinine: random creatinine value (__mean=0.8, sd=0.1__)
    + Symptoms of heart failure: random distribution (probability: __yes = 0.2, no = 0.8__)

```{r data-sim-2,fig.width=7,fig.height=5}

set.seed(5)
n = 3
df_sim_tavi <- r_data_frame(
    n = n,
    age(x = 78:100),
    sex,
    lvef = round(rnorm(n = n, mean = 50, sd = 10), digits = 0),
    crea = round(rnorm(n = n, mean = 1.2, sd = 0.2), digits = 3),
    valid(name = "chf_2w", prob = c(0.1,0.9))
)
df_sim_tavi$Patient = 1:(nrow(df_sim_tavi))
df_sim_tavi$CABG = FALSE
df_sim_tavi$VALVE = "AVR"

df_sim_cabg <- r_data_frame(
    n = n,
    age(x = 40:70),
    sex,
    lvef = round(rnorm(n = n, mean = 50, sd = 10), digits = 0),
    crea = round(rnorm(n = n, mean = 0.8, sd = 0.1), digits = 3),
    valid(name = "chf_2w", prob = c(0.8,0.2))
)
df_sim_cabg$Patient = (nrow(df_sim_tavi)+1):(nrow(df_sim_tavi) + nrow(df_sim_cabg))
df_sim_cabg$CABG = TRUE
df_sim_cabg$VALVE = FALSE

sims <- rbind(df_sim_cabg, df_sim_tavi) %>% group_by(Patient) %>% do(
  calc_sts(proc_valve = .$VALVE, 
           proc_cabg = .$CABG, 
           age = .$Age, 
           gender = .$Sex,
           chf_2w = .$chf_2w,
           crea = .$crea,
           lvef = .$lvef)
  #data_frame(proc_valve = .$VALVE, proc_cabg = .$CABG, age = .$Age, gender = .$Sex, crea = .$crea, lvef = .$lvef)
)

tab_view <- sims %>% 
  dplyr::rename(STS_Risk_Model = Procedure) %>% 
  dplyr::arrange(Patient, Mortality)

knitr::kable(tab_view %>% dplyr::select(Patient, 
                                        Risk_Model = STS_Risk_Model,
                                        Mort = Mortality, 
                                        Mort_Morb = Morbidity_Mortality ,
                                        Stroke = Perm_Stroke, 
                                        Long_Vent = Prolong_Vent, 
                                        ReOP=Reoperation,
                                        Renal_failure,
                                        DSW_Infect=DSW_Infection,
                                        Long_LOS))



# library(dendextend)
# km <- kmeans(dplyr::select(tab_view, -Patient, -STS_Risk_Model) %>% dist, centers = 2)
# km
# dend <- dplyr::select(tab_view, -Patient, -STS_Risk_Model) %>% dist(method = "maximum") %>% hclust(method = "average") %>% as.dendrogram
# dend %>% set("branches_k_color") %>% plot
# dend %>% set("branches_k_color") %>% rect.dendrogram(k=2, border = 8, lty = 5, lwd = 2) %>% plot


p <- radial_plot(sims %>% dplyr::select(-Short_LOS, -Procedure) %>% dplyr::rename(group = Patient) %>% as.data.frame,
                 grid.min = 0, grid.max = 0.3,legend.title = "Patient",
                 plot.extent.x = 1.7, plot.extent.y = 1.2,
                 label.centre.y = FALSE, centre.y = -0.1,
                 group.line.width  = 0.1,
                 group.point.size = 2
                 )
p
```

## STS Score Implemtation

The [STS](http://www.sts.org/)(The Society of Thoracic Surgeons) score is a collection of validated risk models that are regulary updated based on the broad data base of the STS.The [online STS adult cardiac surgery risk calculator](http://riskcalc.sts.org/) is an Angular single page application, which collects the parameters and commincates via JSON to a a web-service("http://riskcalc.sts.org/stswebriskcalc/v1/calculate/stsall ") to obtain the scores. 

The STS score calculation in the riskscorer package is realized by providing a user friendly programmatical interface to the web-service.
The programming interface is able to handle most of data codings automatically. Simple heuristics translate common clinical factor codings. For example "Female", "female" or "f" will all be detected as female gender. Boolean variables such as '0', '1', 'True', 'T', "Y", "Yes" will be detected. Each parameter is described in detail at the corresponding function documentation. Once the score is calculated, a data frame is returned, making it easy to work with the [pipe](https://cran.r-project.org/web/packages/magrittr/index.html) and [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html). 

Terms of use of STS web calculator does not permit the use of the web-service at the moment of writing this vignette. Keep in mind, that this might change and consider to check the terms of use at the [STS score website](http://riskcalc.sts.org/) before using the function.

### Web Service
Thanks to the fantastic [plumber](https://github.com/trestletech/plumber) package, every score calculation function can be easily used as a web service. The hosting of such a service is well documented at the [plumber](http://plumber.trestletech.com/docs/hosting/) documentation.

## ES II implementation

The risk calculation of the EuroScore II is based on a logistic regression model and the coefficients are published in [doi:10.1093/ejcts/ezs043](https://www.ncbi.nlm.nih.gov/pubmed/?term=22378855). Therefor one can easily use the published coefficients and impose a robust R interface over the model.

### ES II examples

```{r}
n = 5

grid_df <- expand.grid(crea = rep(seq(0.8, 3, length.out = n)),
            lvef = rep(seq(10, 60, length.out = n)),
            gender = c("Male", "Female"))

df_sim <- grid_df %>% as.data.frame %>% as.tbl

v_cc_eGFR <- Vectorize(cc_eGFR)

df_sim$age = 75
df_sim$proc_cabg = TRUE
df_sim$proc_valve = TRUE
df_sim$urgency = "elective"
df_sim$height_cm = 170
df_sim$weight_kg = 75
df_sim$sPAP = 33
df_sim$renal = v_cc_eGFR(crea = df_sim$crea, 
                       weight = df_sim$weight_kg, 
                       age = df_sim$age, 
                       sex = df_sim$gender)

sims <- df_sim %>% group_by(lvef, crea, gender) %>% do(
  data_frame(Mortality = calc_esII(valve_surgery = .$proc_valve, 
           CABG = .$proc_cabg, 
           age = .$age, 
           gender = .$gender,
           renal = .$renal,
           sPAP = .$lvef,
           urgency = .$urgency,
           lv = .$lvef))
  #data_frame(proc_valve = .$VALVE, proc_cabg = .$CABG, age = .$Age, gender = .$Sex, crea = .$crea, lvef = .$lvef)
)

#sims_long <- tidyr::gather(data = sims, key = Risk, value = probability, Mortality:Short_LOS)

threejs::scatterplot3js(y = sims$crea, x= sims$lvef, z = sims$Mortality,
                        color = ifelse(sims$gender == "Male", "lightblue", "red"),
                        grid ="TRUE",
                        labels=sprintf("Sex=%s, Crea=%.2f mg/dL, LVEF=%.2f %%, PROM=%.2f %%", 
                                       sims$gender,
                                       sims$crea, 
                                       sims$lvef, 
                                       sims$Mortality*100),
                        renderer = "canvas")

#qplot(x = lvef, y = Mortality, facets = ~gender, color = crea, data = sims, geom = c("point")) + theme_bw()

middle <- function(x) {
  r <- range(x)
  (abs(r[2] - r[1])/2) + r[1]
}

ggplot(data = sims, aes(x=lvef, y=crea, fill=Mortality)) + 
  geom_tile() + theme_minimal() + #geom_point(aes(size = Mortality)) +
  facet_wrap(facets = ~ gender) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = middle(sims$Mortality), 
     space = "Lab", name="EuroScore II\nMortality")

```
