---
title: "DallasLeipzig_DoubleTAVI"
author: "Alexander Meyer"
date: "November 30, 2015"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gtrendsR)
library(riskscorer)

```

## DoubleTAVI Patients STS Score

* Female
* 83 years
* 162 cm, 51 kg
* COPD
* Severe Mitral Regurgitation
* Severe Aortic Stenosis
* Moderate Tricuspid Regurgitation
* NYHA III
* permanent Atrial fibrillation
* Chronic renal failure, Creatinine 1.4 ml/min/1.72m²
* KHK-2
* s. p. PCI RCA + LAD


```{r sim-pat}
#library(riskscorer)
library(dplyr)

n <- 20

age_lvef <- expand.grid(age = seq(from = 60, to = 90, length.out = n),
            lvef = round(seq(from = 60, to = 20, length.out = n)))

case_df <- dplyr::as.tbl(age_lvef)

case_df$group = "Aortic + Mitral"
case_df$gender = "F"
case_df$height_cm = 162
case_df$weight_kg = 51
case_df$proc_cabg = FALSE
case_df$proc_valve = "MVReplacement"
case_df$no_cardiovascular_surgeries = 0
case_df$vd_aortic_stenosis = TRUE
case_df$vd_mitral_regurg = "severe"
case_df$vd_tricuspid_regurg = "moderate"
case_df$crea = 1.4
case_df$chf_2w = TRUE
case_df$chd_known = TRUE
case_df$chd = "KHK-2"
case_df$pre_cv_intervention = TRUE
case_df$pre_pci = TRUE
case_df$pre_pci_interval = ">6h"
case_df$afib = "persistent"

sim_patient <- function(case_desc) {
  case_desc %>% group_by(lvef, age, group) %>% do({
    res <- calc_sts(age = .$age, 
             gender = .$gender,
             weight_kg = .$weight_kg,
             height_cm = .$height_cm,
             proc_cabg = .$proc_cabg, 
             proc_valve = .$proc_valve, 
             no_cardiovascular_surgeries = .$no_cardiovascular_surgeries, 
             vd_aortic_stenosis = .$vd_aortic_stenosis,
             vd_mitral_regurg = .$vd_mitral_regurg,
             vd_tricuspid_regurg = .$vd_tricuspid_regurg,
             crea = .$crea,
             chf_2w = .$chf_2w,
             chd_known = .$chd_known,
             chd = .$chd,
             pre_cv_intervention = .$pre_cv_intervention,
             pre_pci = .$pre_pci,
             pre_pci_interval = .$pre_pci_interval,
             lvef = .$lvef,
             afib = .$afib
             )
    #print(res)
    res
  })
}

sims_full <- sim_patient(case_df)



case_df$group <- "Redo Mitral + Aortic"
case_df$no_cardiovascular_surgeries = 1
sims_redo <- sim_patient(case_df)


case_df$group <- "Isolated AVR"
case_df$proc_valve <- "AVR"
case_df$vd_mitral_regurg <- NULL
case_df$no_cardiovascular_surgeries = 0
sims_no_mvr <- sim_patient(case_df)

sims <- rbind(sims_full, sims_redo, sims_no_mvr)

sims$Mortality <- sims$Mortality * 100

#readr::write_csv(sims, "LeipzigSim_100.csv")
```

## 3D Plots - prohibitive surgical risk?


```{r pressure, echo=FALSE}


# x = seq(1, 10, 1)
# y = seq(1, 10, 1)
# z = outer(x, y, "^")
# 
# persp(x = seq(1, 10, 1), y = seq(1, 10, 1), z, phi = 45, theta = 45,
#   xlab = "X Coordinate (feet)", ylab = "Y Coordinate (feet)",
#   main = "Surface elevation data"
# )
# 
# 
# require(rgl)
# #  open renderer
# open3d()
# #  plot surface
# rgl.surface( 1:10 , 1:10 , runif(100))
# #  Export to png
# rgl.snapshot( "sample.png" , fmt="png", top=TRUE )

# library(animation)
# animation::saveGIF({
#   for (i in seq(from = -60, to = 60, by = 1)) {
#     lattice::wireframe(Mortality ~ lvef*age, data = sims,
#       xlab = "LVEF (%)", ylab = "Age (years)",
#       main = "Mortality",
#       drape = TRUE,
#       colorkey = TRUE,
#       screen = list(z = -60, x = -60)
#     )  
#   }
# }, interval = 0.1, ani.width = 550, ani.height = 550)


library(rgl)
library(dplyr)

sims <- readr::read_csv("ext/LeipzigSim_100.csv")
sims$group <- car::recode(sims$group,
                          "'Redo'='Redo Mitral + Aortic';'Full Surgery'='Aortic + Mitral';'Without Mitral Pathology'='Isolated AVR'")

sims <- sims %>% arrange(desc(group))


colormap <- c("Redo Mitral + Aortic" = "red",
              "Aortic + Mitral" = "orange",
              "Isolated AVR" = "darkgreen")

par3d(cex = 1)
par3d("windowRect"= c(30,30,1024,786))
plot3d(sims$age, #jitter(sims$age, factor = 0.2),
       sims$lvef,
       sims$Mortality,
       col = colormap[sims$group],
#       type = "h",
       xlab = "Age [y]",
       ylab = "LVEF [%]",
       zlab = "Mortality [%]",
       zlim = c(0,20)
       )

#show2d()
#persp3d()

legend3d("right", bty = "n", pch = 1, cex = 1,
         c("Redo Mitral + Aortic", "Aortic + Mitral", "Isolated AVR"),
         col = c("red", "orange", "darkgreen"))
aspect3d(1.5,1,1)
decorate3d(box = FALSE)
grid3d(c("z"))

#play3d()
f <- spin3d(axis = c(0, 0, 1), rpm = 4)
play3d(f)
movie3d(f, duration = 15, fps = 25, movie = "movie3d", clean = FALSE)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
